First and foremost, I am deeply grateful to Jan von Delft for introducing me to the realm of tensor networks and allowing me to pursue this project. His many invitations to present my work highlighted its value, kept me motivated, and taught me the importance of scientific collaboration. From him I have also learned the power of clear communication and teamwork in research.

My heartfelt thanks go to my supervisor, Marc Ritter, whose unwavering support and guidance sustained me throughout this project. Our discussions were always illuminating; through them he showed me what it truly means to be a researcher and provided inspiration that will guide me well into the future. I greatly appreciate his patience and reliability, which allowed me to explore the topic independently while never feeling adrift.

I am indebted to the entire tensor4all team—beginning with Hiroshi Shinaoka, without whom this project would not have existed. Close collaboration with him taught me a great deal about developing computational methods. I am equally grateful to Anna Kauch, Markus Frankenbach, Markus Wallerberger, Samuel Badr, Simone Foderà, and Stefan Rohshap for their insightful conversations and generous assistance.

I also thank everyone at the chair for fostering such a supportive work environment. In particular, Anxiang Ge offered invaluable help during the early stages of this endeavor.

Finally, I extend my deepest gratitude to my friends and family. Their unconditional support has been essential to reaching this milestone.


